---
title: "Predictor Correlations"
author: "Sarah Barnes"
date: "2023-04-15"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

### Install Packages
```{r}
library(ggplot2)
library(GGally)
library(gridExtra)
```

### Load master dataset 

```{r echo=T}
load("master.RData")
```

### --- SCATTER PLOT MATRICES (continuous) ---
```{r, fig.dim=c(12,12)}
cor_matrix_scat <- ggpairs(master[,c(3,8, 18, 20:23, 25:28)]) + theme_bw()
cor_matrix_scat
```

### --- BOXPLOTS FOR CATEGORICAL VS CONTINUOUS ---

#### note that substrate and vegetation can not be done for month because they were only collected in 1 month, and substrate doesn't change by month

#### surface temp, conductivity and clarity can not be done with depth bin because they were only measured at one depth

### --- BOTTOM TYPES VS. DEPTH BINS ---

### Silt
```{r}
silt_dep_box <- ggplot(master, mapping = aes(x = as.factor(depth_bin), y = perc_silt)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "% Silt") +
  annotate("text", x=1.5, y=0.65, size=4, label="Sp. Cor = 0.75" ) +
  theme_bw()
silt_dep_box
```

### Sand
```{r}
sand_dep_box <- ggplot(master, mapping = aes(x = as.factor(depth_bin), y = perc_sand)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "% Sand") +
  annotate("text", x=2, y=0.9, size=4, label="Sp. Cor = 0.26" ) +
  theme_bw()
sand_dep_box
```

### Clay
```{r}
clay_dep_box <- ggplot(master, mapping = aes(x = as.factor(depth_bin), y = perc_clay)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "% Clay") +
  annotate("text", x=2.5, y=0.75, size=4, label="Sp. Cor = -0.71" ) +
  theme_bw()
clay_dep_box
```

### Rock
```{r}
rock_dep_box <- ggplot(master, mapping = aes(x = as.factor(depth_bin), y = perc_rock)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "% Rock") +
  annotate("text", x=1, y=0.15, size=4, label="Sp. Cor = 0.76" ) +
  theme_bw()
rock_dep_box
```

### --- VEGETATION VS. DEPTH BINS ---

### Vegetation Height
```{r}
veg_height_box <- ggplot(subset(master, depth_bin != 0), mapping = aes(x = as.factor(depth_bin), y = avg_vegheight)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "Average Vegetation Height (m)") +
  annotate("text", x=2.5, y=0.4, size=4, label="Sp. Cor = -0.43" ) +
    theme_bw()
veg_height_box
```

### Vegetation Cover
```{r}
veg_cover_box <- ggplot(subset(master, depth_bin != 0), mapping = aes(x = as.factor(depth_bin), y = avg_vegcover)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "Average % Cover") +
  annotate("text", x=2.5, y=60, size=4, label="Sp. Cor = -0.49" ) +
  theme_bw()
veg_cover_box
```

### --- SURFACE TEMP VS. MONTH ---
```{r}
srf_temp_box <- ggplot(master, mapping = aes(x = as.factor(month), y = surface_temp)) + 
  geom_boxplot() +
  labs( x= "Month", y= "Surface Temp") +
  annotate("text", x=1.5, y=25, size=4, label="Sp. Cor = 0.92" ) +
  theme_bw()
srf_temp_box
```

### --- CLARITY VS. MONTH ---
```{r}
clar_box <- ggplot(master, mapping = aes(x = as.factor(month), y = clarity)) + 
  geom_boxplot() +
  labs( x= "Month", y= "Clarity") +
  annotate("text", x=2.5, y=8, size=4, label="Sp. Cor = 0.33" ) +
  theme_bw()
clar_box
```

### --- CONDUCTIVITY VS. MONTH ---
```{r}
cond_box <- ggplot(master, mapping = aes(x = as.factor(month), y = conductivity)) + 
  geom_boxplot() +
  labs( x= "Month", y= "Conductivity") +
  annotate("text", x=1.5, y=8.4, size=4, label="Sp. Cor = 0.91" ) +
  theme_bw()
cond_box
```

### --- TEMPERATURE VS. MONTH ---

#### note here that there are 6 less deep sets because the temperatures are cut based on net sets 
and we only set the net at 30 m 3x instead of 9x in the late July set
```{r}
temp_avg_mon_box <- ggplot(master, mapping = aes(x = as.factor(month), y = temp_avg)) + 
  geom_boxplot() +
  labs( x= "Month", y= "Average Logger Temperature") +
  annotate("text", x=1.5, y=22, size=4, label="Sp. Cor = 0.84" ) +
  theme_bw()
temp_avg_mon_box
```

### --- TEMPERATURE VS. DEPTH BIN ---
```{r}
temp_avg_dep_box <- ggplot(master, mapping = aes(x = as.factor(depth_bin), y = temp_avg)) + 
  geom_boxplot() +
  labs( x= "Depth Bin", y= "Average Logger Temperature") +
  annotate("text", x=3.5, y=25, size=4, label="Sp. Cor = -0.73" ) +
  theme_bw()
temp_avg_dep_box
```

### --- ALL TOGETHER ---

```{r, echo=F, fig.dim=c(10,10)}
# make a blank for aesthitics
blank <- ggplot() +
  geom_blank()

comp_box_cor <- grid.arrange(clay_dep_box, sand_dep_box, silt_dep_box, rock_dep_box, 
                             veg_height_box, veg_cover_box,temp_avg_dep_box, blank, srf_temp_box, clar_box, cond_box,
                             temp_avg_mon_box,
                             nrow=3, ncol=4)
comp_box_cor
```